---
title: "Chapter 20: K-means Clustering"
output: html_notebook
---

__Note__: Some results may differ from the hard copy book due to the changing of
sampling procedures introduced in R 3.6.0. See http://bit.ly/35D1SW7 for more
details. Access and run the source code for this notebook [here](https://rstudio.cloud/project/801185).

Hidden chapter requirements used in the book to set the plotting theme and load
packages used in hidden code chunks:

```{r setup}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  cache = FALSE
)

# Set the graphical theme
ggplot2::theme_set(ggplot2::theme_light())

# Load required packages
library(tidyr)
library(purrr)
```

## Prerequisites

For this chapter we’ll use the following packages:

```{r kmeans-pkgs}
# Helper packages
library(dplyr)       # for data manipulation
library(ggplot2)     # for data visualization
library(stringr)     # for string functionality

# Modeling packages
library(cluster)     # for general clustering algorithms
library(factoextra)  # for visualizing cluster results
```

To illustrate _k_-means concepts we'll use the `mnist` and `my_basket` data sets:

```{r kmeans-data}
mnist <- dslabs::read_mnist()

url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"
my_basket <- readr::read_csv(url)
```

## Distance measures

Figure 20.1:

```{r correlation-distance-example, fig.width=7, fig.cap='Correlation-based distance measures will capture the correlation between two observations better than a non-correlation-based distance measure; regardless of magnitude differences.'}
# generate data
corr_ex <- tibble(
  v = 1:20,
  obs_1 = sample(5:7, 20, replace = TRUE),
  obs_2 = sample(4:10, 20, replace = TRUE)
) %>%
  mutate(obs_3 = obs_2 * 2 + sample(0:1, 1))

corr_ex %>%
  gather(Observation, value, obs_1:obs_3) %>%
  ggplot(aes(v, value, color = Observation)) +
  geom_line(size = 1) +
  scale_colour_manual(values = c("#00AFBB", "#E7B800", "#FC4E07")) +
  scale_x_continuous("Variable index") +
  scale_y_continuous("Some arbitrary measure")
```


## Defining clusters

Figure 20.2:

```{r kmeans-clusters-good-better-best, fig.height=3.5, fig.width=10, fig.cap="Total within-cluster variation captures the total distances between a cluster's centroid and the individual observations assigned to that cluster. The more compact the these distances, the more defined and isolated the clusters are."}
# Generate data
create_data <- function(sd) {
  data_frame(
    x1 = c(rnorm(100, sd = sd), rnorm(100, sd = sd) + 3),
    x2 = c(rnorm(100, sd = sd), rnorm(100, sd = sd) - 2)
  ) %>%
    mutate(`W(Ck)` = case_when(
      sd == 0.5  ~ "Best",
      sd == 0.75 ~ "Better",
      sd == 1   ~ "Good"
    ))
}
df <- map(c(0.5, 0.75, 1), create_data)

# Compute and add cluster info to data
k2 <- map(df, ~ kmeans(.x[, 1:2], 2, nstart = 20))
df <- map2(df, k2, ~ mutate(.x, cluster = .y$cluster)) %>%
  map2_dfr(k2, ~ inner_join(.x, .y$centers %>% 
                          as.data.frame() %>% 
                          mutate(cluster = row_number()), by = "cluster")
       ) %>%
  rename(x1 = x1.x, x2 = x2.x, x_center = x1.y, y_center = x2.y) %>%
  mutate(`W(Ck)` = factor(`W(Ck)`, levels = c("Good", "Better", "Best")))

# Plot results
df %>%
  ggplot(aes(colour = factor(cluster))) +
  facet_wrap(~ `W(Ck)`) +
  geom_segment(aes(x = x1, xend = x_center, y = x2, yend = y_center), lty = "dashed", alpha = .5) +
  geom_point(aes(x_center, y_center), size = 4) +
  geom_point(aes(x1, x2), show.legend = FALSE, alpha = .5) +
  scale_x_continuous(bquote(X[1]), breaks = NULL, labels = NULL) +
  scale_y_continuous(bquote(X[2]), breaks = NULL, labels = NULL) +
  theme(legend.position = "none")
```

Figure 20.3:

```{r non-linear-boundaries, fig.cap='The assumptions of k-means lends it ineffective in capturing complex geometric groupings; however, spectral clustering allows you to cluster data that is connected but not necessarily clustered within convex boundaries.'}
# Generate data
set.seed(111)
obj <- mlbench::mlbench.spirals(200, 1, 0.025)
df <- data.frame(
  x = obj$x[, 1],
  y = obj$x[, 2],
  class = obj$classes
)

# Plot data
p1 <- ggplot(df, aes(x, y)) +
  geom_point() +
  xlab(NULL) +
  ylab(NULL) +
  ggtitle('(A) Original spiral data')

# Run k-means
kmeans_on_spiral <- kmeans(df[, 1:2], 2)
df$kmeans_clusters <- kmeans_on_spiral$cluster
p2 <- ggplot(df, aes(x, y, color = kmeans_clusters)) +
  geom_point(show.legend = FALSE) +
  xlab(NULL) +
  ylab(NULL) +
  ggtitle('(B) k-means clusters')

# Plot results
sc <- kernlab::specc(as.matrix(df[, 1:2]), centers = 2)
df$spec_clusters <- sc@.Data
p3 <- ggplot(df, aes(x, y, color = spec_clusters)) +
  geom_point(show.legend = FALSE) +
  xlab(NULL) +
  ylab(NULL) +
  ggtitle('(C) Spectral clusters')

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```


## _k_-means algorithm

Figure 20.4:

```{r random-starts, fig.height=6, fig.width=10, fig.cap='Each application of the k-means algorithm can achieve slight differences in the final results based on the random start.'}
# Generate data
df <- data_frame(
    x1 = c(rnorm(100), rnorm(100) + 3),
    x2 = c(rnorm(100), rnorm(100) - 2)
)

# Compute and plot results
map(1:6, ~ kmeans(df, 3)) %>%
  map2_dfr(1:6, ~ df %>% mutate(
    cluster = .x$cluster,
    name = paste0("Iteration: ", .y, ";  W(Ck): ", round(.x$tot.withinss, 2))
    )) %>%
  ggplot(aes(x1, x2, colour = cluster)) +
  geom_point(show.legend = FALSE, size = 1) +
  facet_wrap(~ name, nrow = 2)
```

## Clustering digits

```{r mnist-kmeans}
features <- mnist$train$images

# Use k-means model with 10 centers and 10 random starts
mnist_clustering <- kmeans(features, centers = 10, nstart = 10)

# Print contents of the model output
str(mnist_clustering)
```

```{r plot-kmeans-mnist-centers, fig.height=4, fig.width=12, fig.cap='Cluster centers for the 10 clusters identified in the MNIST training data.'}
# Extract cluster centers
mnist_centers <- mnist_clustering$centers

# Plot typical cluster digits
par(mfrow = c(2, 5), mar = c(0.5, 0.5, 0.5, 0.5))
layout(matrix(seq_len(nrow(mnist_centers)), 2, 5, byrow = FALSE))
for (i in seq_len(nrow(mnist_centers))) {
  image(matrix(mnist_centers[i, ], 28, 28)[, 28:1], 
        col = gray.colors(12, rev = TRUE), xaxt = "n", yaxt = "n")
}
```

```{r mnist-clustering-confusion-matrix, fig.cap='Confusion matrix illustrating how the k-means algorithm clustered the digits (x-axis) and the actual labels (y-axis).'}
# Create mode function
mode_fun <- function(x){  
  which.max(tabulate(x))
}

mnist_comparison <- data.frame(
  cluster = mnist_clustering$cluster,
  actual = mnist$train$labels
) %>%
  group_by(cluster) %>%
  mutate(mode = mode_fun(actual)) %>%
  ungroup() %>%
  mutate_all(factor, levels = 0:9)

# Create confusion matrix and plot results
yardstick::conf_mat(
  mnist_comparison, 
  truth = actual, 
  estimate = mode
) %>%
  autoplot(type = 'heatmap')
```

## How many clusters? 

```{r elbow-method, fig.cap="Using the elbow method to identify the preferred number of clusters in the my basket data set."}
fviz_nbclust(
  my_basket, 
  kmeans, 
  k.max = 25,
  method = "wss",
  diss = get_dist(my_basket, method = "spearman")
)
```

## Clustering with mixed data

```{r}
# Full ames data set --> recode ordinal variables to numeric
ames_full <- AmesHousing::make_ames() %>%
  mutate_if(str_detect(names(.), 'Qual|Cond|QC|Qu'), as.numeric)

# One-hot encode --> retain only the features and not sale price
full_rank  <- caret::dummyVars(Sale_Price ~ ., data = ames_full, 
                               fullRank = TRUE)
ames_1hot <- predict(full_rank, ames_full)

# Scale data
ames_1hot_scaled <- scale(ames_1hot)

# New dimensions
dim(ames_1hot_scaled)
```

```{r kmeans-silhouette-mixed, fig.width=7, fig.height=4, fig.cap="Suggested number of clusters for one-hot encoded Ames data using k-means clustering and the elbow criterion."}
set.seed(123)

fviz_nbclust(
  ames_1hot_scaled, 
  kmeans, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r}
# Original data minus Sale_Price
ames_full <- AmesHousing::make_ames() %>% select(-Sale_Price)

# Compute Gower distance for original data
gower_dst <- daisy(ames_full, metric = "gower")
```

```{r gower-based-clustering, eval=FALSE}
# You can supply the Gower distance matrix to several clustering algos
pam_gower <- pam(x = gower_dst, k = 8, diss = TRUE)
diana_gower <- diana(x = gower_dst, diss = TRUE)
agnes_gower <- agnes(x = gower_dst, diss = TRUE)
```

## Alternative partitioning methods

```{r pam, fig.width=7, fig.height=4, fig.cap="Total within sum of squares for 1-25 clusters using PAM clustering."}
fviz_nbclust(
  ames_1hot_scaled, 
  pam, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r clara}
# k-means computation time on MNIST data
system.time(kmeans(features, centers = 10))

# CLARA computation time on MNIST data
system.time(clara(features, k = 10))
```